### 2016-02-05 Posters.tex
corrections files added

% !TEX encoding = UTF-8 Unicode
% !TEX program = pdflatex

\chapter{List of Posters presented}\label{posterslist}

\begin{sloppypar}
All posters can be seen in the digital poster exhibition on the Eagle website.
\url{www.eagle-network.eu/about/events/eagle2016/digital-poster-exhibition/}

\end{sloppypar}

\begin{table}
{\small
\addtolength{\tabcolsep}{-0.8mm}

\begin{tabular*}{\textwidth}{p{4.5cm}|p{6cm}}
%    \rowcolor{black!70}
%    \multicolumn{1}{c}{\textcolor{white}{\textbf{Author}}} & 
%    \multicolumn{1}{c}{\textcolor{white}{\textbf{Title}}} 
%    \tabularnewline
%    
%    \endfirsthead
%    \multicolumn{2}{c}
%    {{\bfseries \tablename\ \thetable{} -- continued from previous page}} \tabularnewline
%    \rowcolor{black!70}
%    \multicolumn{1}{c}{\textcolor{white}{\textbf{Author}}} &
%    \multicolumn{1}{c}{\textcolor{white}{\textbf{Title}}}
%    \tabularnewline
%    
%    \endhead
% \rowcolor{black!70}
%    \multicolumn{2}{r}{{ \textcolor{white}{-- continued on next page}}} \tabularnewline 
%
%    \endfoot
%    
%    \endlastfoot
%   
\toprule Jan-Mathieu Carbon, Saskia Peels and Vinciane Pirenne-Delforge & The Collection of Greek Ritual Norms Project (CGRN) \\\hline
Stefano Caneva and Guido D'Ippolito & Pasquino: Talking Statues and Inscriptions in Rome\\\hline
Mat Carbon, Mario Paganini and Stella Skaltsa & The Copenhagen Associations Project (CAP)\\\hline
Emilia Mataix Ferrandiz & Bridging the gap Linking commercial epigraphy and Roman law\\\hline
Barbara Mazzei and Agnese Pergola & The PCAS Photographic Digital Archive: Fruition of Epigraphic Images\\\hline
Barbara De Nicolò & “Talking Stones”: from Stone to History. A Latin Epigraphy workshop for Primary School\\\hline
Sara Chiarini and Kirsten Jahn & TheDeMa (Thesaurus Defixionum Magdeburgensis) – A digital companion to ancient curse tablets\\\hline
Daniele Silvi and Fabio Ciotti & Thematic annotation and ontologies: application to digital epigraphy\\\hline
Loredana Francesca Tedeschi & Integration of geomatic techniques for the analysis and geolocation of graphic forms of the Filocalian letter in the Elogia Martyrum by Pope Damasus(366-384)\\\hline
Fulvia Mainardis, Claudio Zaccaria, Andrea Favretto and Giovanni Mauro & A WebGIS for the knowledge of the epigraphic locations in the Friuli Venezia Giulia region (North-East of Italy)\\\hline
Petra Janouchova & INSCRIPTIONS TO GO: Mobile platform for offline epigraphic data collection\\\hline
Tamar Kalkhitashvili and Eka Kvirkvelia & Epigraphic Corpus of Georgia\\\hline
Gaia Trombin and Antonio Bonaldo & Out of the wall. 3D documentation and restoration of an endangered Roman altar\\\hline
Gaia Lembi & Saxa Judaica: Delving Digitally into Jewish Inscriptions\\\hline
Olga Pelcer-Vujacic. Epigraphica Montenegrina & The road so far and the journey ahead\\\hline
Marion Lamé, Federico Ponchio, Ivan Radman-Livaja and Bruce Robertson & Teaching (Digital) Epigraphy\\\hline
Claudio Prandoni, Antonella Fresa & Civic Epistemologies \\\hline
Silvia Marchesini & Ancient European Languages \\
Antonella Ferraro & From Paper to Web \\
Lavinio Del Monaco & Le iscrizioni di Palestrina \\
\bottomrule

\end{tabular*}}

\caption{Posters presented at the conference}
\label{tab:table2}
\end{table}